Modal type theory based on the intuitionistic epistemic logic

### 2018-01-18 article.tex
added

  5) $M \rightarrow_{\beta} N \Rightarrow {\bf pure \: M} \rightarrow_{\beta} {\bf pure \: N}$

  6) $\lambda x. f x \rightarrow_{\eta} f$;

  7) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;

  10) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \rightarrow_{\eta} {\bf pure \:} N$;

  11) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;

  12) $M \rightarrow_{\beta} N \Rightarrow {\bf pure \: M} \rightarrow_{\eta} {\bf pure \: N}$
  \end{defin}

  Let us show the next simple observation, which immeadelty follows from the previous definition.

    If $M \rightarrow_{\beta \eta} N$, then ${\bf pure \:} M \rightarrow_{\beta \eta} {\bf pure \:} N$.
  \end{lemma}

@@

1) $[\![\Gamma \vdash {\bf pure \:} ((\lambda x. M) N) : {\bf K}B]\!] = [\![\Gamma \vdash {\bf pure \:} (M [x := N]]) : {\bf K}B]\!]$

$\begin{array}{lll}
&[\![\Gamma \vdash {\bf pure \:} (\lambda x. M) N : {\bf K}B]\!] = & \text{By interpretation} \\
&\eta_{[\![B]\!]} \circ (\epsilon \circ \langle \Lambda([\![M]\!]), [\![N]\!]\rangle) =& \text{Property of $\times$} \\
&\eta_{[\![B]\!]} \circ (\epsilon \circ (\Lambda([\![M]\!]) \times id_{[\![A]\!]}) \circ \langle id_{[\![\Gamma]\!]}, [\![N]\!] \rangle) = & \text{Associativity of composition} \\
&\eta_{[\![B]\!]} \circ ((\epsilon \circ (\Lambda([\![M]\!]) \times id_{[\![A]\!]})) \circ \langle id_{[\![\Gamma]\!]}, [\![N]\!] \rangle) = & \text{Exponentiation property} \\
&\eta_{[\![B]\!]} \circ ([\![M]\!] \circ \langle id_{[\![\Gamma]\!]}, [\![N]\!] \rangle) = & \text{Substitution lemma} \\
&\eta_{[\![B]\!]} \circ [\![M [\vec{x}, x := \vec{x}, N]]\!] =& \text{By interpretation} \\
&[\![\Gamma \vdash {\bf pure \:} (M [x := N]]) : {\bf K}B]\!]&
\end{array}$

@@

2) $[\![\Gamma \vdash {\bf pure \:} (\pi_i \langle [\![M_1]\!], [\![M_2]\!] \rangle) : {\bf K}A_i ]\!] = [\![\Gamma \vdash {\bf \: pure} M_i : {\bf K}A_i ]\!]$

@@

$\begin{array}{lll}
&[\![\Gamma \vdash {\bf pure \:} (\pi_i \langle M_1, M_2] \rangle) : {\bf K}A_i ]\!] = & \text{By interpretation} \\
&\eta_{[\![A_i]\!]} \circ \pi_i \circ \langle [\![M_1]\!], [\![M_2]\!] \rangle = & \text{Property of $\times$} \\
&\eta_{[\![A_i]\!]} \circ [\![M_i]\!] =& \text{By interpetation} \\
&[\![\Gamma \vdash {\bf pure \:} M_i : {\bf K}A_i]\!]&
\end{array}$

@@

1) $[\![\Gamma \vdash {\bf pure \:} (\lambda x. M x) : {\bf K}(A \to B)]\!] = [\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \to B)]\!]$.

$\begin{array}{lll}
&[\![\Gamma \vdash {\bf pure \:} (\lambda x. M x) : {\bf K}(A \to B)]\!] = & \text{By interpetation} \\
&\eta_{[\![B]\!]^{[\![A]\!]}} \circ \Lambda(\epsilon \circ [\![M]\!] \times id_{[\![A]\!])}& \text{Exponentiation property} \\
&\eta_{[\![B]\!]^{[\![A]\!]}} \circ [\![M]\!] = & \text{By interpretation} \\
&[\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \to B)]\!]&
\end{array}$

2) $[\![\Gamma \vdash {\bf pure \:} \langle \pi_1 M, \pi_2 M \rangle : {\bf K}(A \times B)]\!] = [\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \times B)]\!]$

$\begin{array}{lll}
&[\![\Gamma \vdash {\bf pure \:} \langle \pi_1 M, \pi_2 M \rangle : {\bf K}(A \times B)]\!] = & \text{By interpetation} \\
&\eta_{[\![A]\!] \times [\![B]\!]} \circ \langle \pi_1 \circ [\![M]\!], \pi_2 \circ [\![M]\!] \rangle  = & \text{By the property of a product of morphisms} \\
&\eta_{[\![A]\!] \times [\![B]\!]} \circ [\![M]\!] = & \text{By interpetation} \\
&[\![\Gamma \vdash {\bf pure \:} M : {\bf K}(A \times B)]\!]&
\end{array}$

3) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = [\![\vdash {\bf pure \:} N : {\bf K A}]\!]$.

@@

@@

\begin{proof}
$ $
Straightforwardly follows from two previous lemmas.
\end{proof}

>>>>

  5) ${\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N \rightarrow_{\beta} {\bf pure \:} N [\vec{x} := \vec{M}]$

  6) $M \rightarrow_{\beta} N \Rightarrow {\bf pure \: M} \rightarrow_{\beta} {\bf pure \: N}$

  7) $\lambda x. f x \rightarrow_{\eta} f$;

  8) $\langle \pi_1 P, \pi_2 P \rangle \rightarrow_{\eta} P$;

  9) ${\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N \rightarrow_{\eta} {\bf pure \:} N$;

  10) ${\bf let \: pure \:} x = M {\: \bf in \: } x \rightarrow_{\eta} M$;

  11) $M \rightarrow_{\eta} N \Rightarrow {\bf pure \: M} \rightarrow_{\eta} {\bf pure \: N}$
  \end{defin}

@@

1) $[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B]\!] = [\![\Gamma \vdash {\bf pure \:} N [\vec{x} := \vec{M}] : {\bf K}B]\!]$

@@

$\begin{array}{lll}
&[\![\Gamma \vdash {\bf let \: pure \:} \vec{x} = {\bf pure \:} \vec{M} {\: \bf in \:} N : {\bf K}B]\!] = & \\
&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By interpretation}& \\
&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ \langle \eta_{[\![A_1]\!]} \circ [\![M_1]\!],\dots,\eta_{[\![A_n]\!]} \circ [\![M_n]\!] \rangle = \\
&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By the property of a pair of morphisms}& \\
&\mathcal{K}([\![N]\!]) \circ \ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ (\eta_{[\![A_1]\!]} \times \dots \times \eta_{[\![A_n]\!]}) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]\rangle =& \\
&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Associativity of composition}& \\
&\mathcal{K}([\![N]\!]) \circ (\ast_{[\![A_1]\!],\dots,[\![A_n]\!]} \circ (\eta_{[\![A_1]\!]} \times \dots \eta_{[\![A_n]\!]})) \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By the definition of an applicative functor}& \\
&\mathcal{K}([\![N]\!]) \circ \eta_{[\![A_1]\!] \times \dots \times [\![A_n]\!]} \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Naturality of $\eta$}& \\
&\eta_{[\![B]\!]} \circ [\![N]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!] \rangle =& \\
&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Associativity of composition}& \\
&\eta_{[\![B]\!]} \circ ([\![N]\!] \circ \langle [\![M_1]\!], \dots, [\![M_n]\!]) \rangle =& \\
&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad Simultaneous substitution lemma}& \\
&\eta_{[\![B]\!]} \circ [\![N [\vec{x} := \vec{M}]]\!]& \\
&\text{\quad\quad\quad\quad\quad\quad\quad\quad\quad By interpetation}& \\
&[\![\Gamma \vdash {\bf pure \:} (N [\vec{x} := \vec{M}]) : {\bf K}B]\!]&
\end{array}$

2)

If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.

@@

If $\Gamma \vdash M : A$ and $M \rightarrow_{\beta} N$, then $\Gamma \vdash N : A$ by subject reduction.

By assumption $[\![\Gamma \vdash M : A]\!] = [\![\Gamma \vdash N : A]\!]$.

So $\eta_{[\![A]\!]} \circ [\![\Gamma \vdash M : A]\!] = \eta_{[\![A]\!]} \circ [\![\Gamma \vdash N : A]\!]$.

Hence $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.

@@

1) $[\![\vdash {\bf let \: pure \:} \underline{\quad} = \underline{\quad} {\: \bf in \:} N : {\bf }K A ]\!] = [\![\vdash {\bf pure \:} N : {\bf K A}]\!]$.

@@

2)

If $\Gamma \vdash M : A$ and $M \rightarrow_{\eta} N$, then $[\![\Gamma \vdash {\bf pure \:} M : {\bf K}A]\!] = [\![\Gamma \vdash {\bf pure \:} N : {\bf K}A]\!]$.

Similar to case with $\beta$-reduction.

@@

\subsection{Completeness}
